Embedding Intuitionistic into Classical Logic

### 2022-07-12 CSL/lipics-v2021-EIICL.tex
typos

\begin{theorem}\label{fullFOtranslation}
	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$. For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$. Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal S^\#$ by including formulas as follows:
	\begin{itemize}
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\to (B_\psi^\#(\vec b_\psi, u)\circ C_\psi^\#(\vec c_\psi, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\circ C_\psi(\vec c_\psi)))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\circ B_\psi^\#(\vec b_\psi, u))\to C_\psi^\#(\vec c_\psi, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\circ B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to(A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u))\to B_\psi^\#(\vec b_\psi, f_\psi(\vec z_\psi, u)))\to C_\psi^\#(\vec c_\psi, u))$ for each $\psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
		\item  $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \forall x(E(x, f_\psi(\vec z_\psi, u))\to A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u)))\to B^\#(\vec b_\psi, u))$\\for each $\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec t_B))\in\mathcal S$.
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec a_\psi, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec t(A_\psi(\vec a_\psi)\to \forall xB_\psi(\vec b_\psi))\in\mathcal S$.
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec z_\psi, u))\to B^\#_\psi(\vec b_\psi))$\\for each $\forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec b_\psi)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec z_\psi(A_\psi(\vec a_\psi)\to \exists xB_\psi(\vec b_\psi))\in\mathcal S$
	\end{itemize}
	Then $\bigwedge S\to P$ is intuitionistically iff $K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol and
	\begin{align*}
		K := &\exists xE(x, s)\wedge \forall x\forall y\forall z(\neg(E(x, z)\wedge E(z, y)))\wedge\\
		& \bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
		& \bigwedge\{\forall\vec z\forall u(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is an $n$-ary function symbol in original signature}\}
	\end{align*}
	The size of the obtained formula is linear in the input. However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
\end{theorem}

>>>>

\begin{theorem}\label{fullFOtranslation}
	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$. For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$. Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal S^\#$ by including formulas as follows:
	\begin{itemize}
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\to (B_\psi^\#(\vec b_\psi, u)\circ C_\psi^\#(\vec c_\psi, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\circ C_\psi(\vec c_\psi)))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\circ B_\psi^\#(\vec b_\psi, u))\to C_\psi^\#(\vec c_\psi, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\circ B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to(A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u))\to B_\psi^\#(\vec b_\psi, f_\psi(\vec z_\psi, u)))\to C_\psi^\#(\vec c_\psi, u))$ for each $\psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
		\item  $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \forall x(E(x, f_\psi(\vec z_\psi, u))\to A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u)))\to B^\#(\vec b_\psi, u))$\\for each $\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec t_B))\in\mathcal S$.
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec a_\psi, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec t(A_\psi(\vec a_\psi)\to \forall xB_\psi(\vec b_\psi))\in\mathcal S$.
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec z_\psi, u))\to B^\#_\psi(\vec b_\psi))$\\for each $\forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec b_\psi)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec z_\psi(A_\psi(\vec a_\psi)\to \exists xB_\psi(\vec b_\psi))\in\mathcal S$
	\end{itemize}
	Then $\bigwedge S\to P$ is intuitionistically iff $\exists xE(x, s)\to K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol, $\Sigma$ is the original signature and
	\begin{align*}
		K := &\:\bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}
	\end{align*}
	The size of the obtained formula is linear in the input. However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
\end{theorem}
